Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-06 paper/paper.tex
anglican vs clojure

\subsection*{Contributions} This paper brings the following
major contributions:
\begin{itemize}
	\item Design and implementation of a probabilistic
		programming language Anglican involving tight bilateral
		integration with a general-purpose programming language.
	\item Definition of semantics of probabilistic forms in
		Anglican which facilitates full separation of
		probabilistic programs from inference algorithms.
	\item Stateless representation of random processes, seamlessly integrated
		into pure functional computation.
	\item Techniques for efficient and compact implementation of
		inference algorithms, such as representation of
		inference results as a lazy sequence of samples and 
		a novel scheme for addressing of checkpoints.
\end{itemize}

@@

>>>>

\subsection*{Contributions} This paper brings the following
major contributions:
\begin{itemize}
    \item Design and implementation of a probabilistic programming language
        Anglican involving tight bilateral integration with a general-purpose
        programming language.
    \item Semantics of probabilistic forms in Anglican which facilitates full
        separation of probabilistic programs from inference algorithms.
    \item Stateless representation of random processes, seamlessly integrated
        into pure functional computation.
    \item Techniques for efficient and compact implementation of inference
        algorithms, such as representation of inference results as a lazy
        sequence of samples and a novel scheme for addressing of checkpoints.
\end{itemize}

@@

To justify the claim that Anglican is an \textit{efficient}
implementation, performance of Anglican programs must evaluated
both against Clojure programs and against inference tasks in
other probabilistic programming languages and environments.
Anglican passes two additional parameters to every function, adds
extra code to variable bindings, and passes every function
call through the trampoline. Hence difference in performance
between Anglican and Clojure should be most noticeable on
programs which involve many function calls. One such example is
the famous \textit{Towers of Hanoi}:

\begin{lstlisting}[style=default]
(defn towers-of-hanoi [n from to via]
  (when (not= n 1)
      (towers-of-hanoi (dec n) from via to)
      (towers-of-hanoi (dec n) via to from)))
\end{lstlisting}

We run this program for n equal to 10, 15, 20, and 25. Clojure
is consistently only twice as fast as Anglican
(Table~\ref{tbl:hanoi}); considering the amount of overhead
Anglican introduces, this is a good result.
\begin{table}
    \centering
    \begin{tabular}{r|r|r|r|r}
        {\bf n}  & {\bf 10} & {\bf 15} & {\bf 20} &{\bf 25} \\ \hline
        Clojure &  0.32 & 1.45 & 38.1  &  1245.9 \\ \hline
        Anglican & 0.63 & 3.00 & 75.9 & 2381.1 
    \end{tabular}
    \caption{Towers of Hanoi. Average running times, in
    seconds, of Clojure and Anglican, averaged over 100 runs.
    Clojure is only twice as fast as Anglican.}
    \label{tbl:hanoi}
\end{table}
